R for Clinical Study Reports and Submission

### 2023-07-16 tlf-baseline.qmd
improve `tlf-baseline.qmd`

In this chapter, we illustrate how to create a simplified
baseline characteristics table for a study.

@@

```{r}
library(table1)
library(r2rtf)
library(haven)
library(dplyr)
library(tidyr)
library(stringr)
library(tools)
```

As in previous chapters, we first read the `adsl` dataset that contains all
the required information for the baseline characteristics table.

@@

For simplicity, we only analyze `SEX`, `AGE` and, `RACE` in this example
using the table1 R package. More details of the table1 R package can
be found in the package
[vignettes](https://benjaminrich.github.io/table1/vignettes/table1-examples.html).

The table1 R package directly creates an HTML report.

@@

The code below transfer the output into a dataframe
that only contains ASCII characters
recommended by regulatory agencies.
`tbl_base` is used as input for r2rtf to create the final report.

@@

We define the format of the output. We highlight items that
are not discussed in previous discussion.

`text_indent_first` and `text_indent_left` are used to control the indent
space of text. They are helpful when you need to control the white space
of a long phrase, "AMERICAN INDIAN OR ALASKA NATIVE"
in the table provides an example.

@@

In conclusion, the procedure to generate demographic and baseline characteristics table is summarized as follows:

- Step 1: Read the data set.
- Step 2: Use `table1::table1()` to get the baseline characteristics table.
- Step 3: Transfer the output from Step 2 into a data frame that only contains ASCII characters.
- Step 4: Define the format of the RTF table by using the R package r2rtf.

>>>>

This chapter provides a step-by-step guide on creating a simple 
baseline characteristics table in RTF format.
The expected output of the table is as follows:

@@

```{r}
library(haven)
library(dplyr)
library(tidyr)
library(stringr)
library(tools)
library(table1)
library(r2rtf)
```

## Input data

We begin by reading the `adsl` dataset, 
which comprises all the necessary information for generating the baseline characteristics table.

@@

For simplicity, we only analyze `SEX`, `AGE` and, `RACE` in this example.

## Output data

The table1 R package directly creates an HTML report for the baseline characteristic table.
More details of the table1 R package can
be found in the package
[vignettes](https://benjaminrich.github.io/table1/vignettes/table1-examples.html).

@@

let's transfer the output into a dataframe that 
exclusively consists of ASCII characters, as recommended by regulatory agencies.

@@

## Reporting

By using `tbl_base` as the input, we can generate the baseline characteristics table.

In the code below, `text_indent_first` and `text_indent_left`
are used to control the indentation space of the text. 
These parameters come in handy when you need to manage the whitespace within a lengthy phrase.

@@
